Summary Tables

### 2023-11-29 summary_tables.Rmd
Commented out health related outputs and kept only the necessary bits re distance, duration and definition of scenario

# Distance categories

@@

### Baseline mode split for all cities (%)
```{r}
# create baseline mode share table containing information for all cities

# prep data
trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
  mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>%
  janitor::adorn_totals('row')

# print table to html
kable(trip_prop, caption = 'Baseline mode share (%) for all cities - only selected modes')

```{r unload_tidyverse, echo = F, message = F}
suppressWarnings({
  detach("package:tidyverse", character.only = T)
})

@@

### Distance table showing average baseline distances in km travelled by each person in the synthetic population by mode for all cities

```{r}
# create table showing average baseline distances by mode for all cities
trip_dist <- trip_dist %>% janitor::adorn_totals('row')

kable(trip_dist, caption = "Avg. baseline distance (km) per person in the synthetic population per day across all cities")

## Distance tables scaled by using total model age population (per year in km)

City specific distance tables for baseline and scenarios for the entire population with ages considered in the model

```{r trip_tot_dist = "asis", message=F}
# create distance tables for the yearly distances travelled by the entire population
bl_td <- list()

  # extract daily population distances for ages considered in model and multiply by 365 to to get yearly distances
  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
    mutate_if(is.numeric, list(~round((.) * 365, 3)))

  # update scenario names
  colnames(td)[2:ncol(td)] <- scen_names_updated

  # initialise bl_td with stage_mode
  if (length(bl_td) == 0){
    bl_td <- td %>% dplyr::select(stage_mode)
  }

  # calculate row and column totals
  td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))

  # extract baseline distances
  td1 <- td %>% dplyr::select(stage_mode, Baseline)

  # join baseline distances for all cities
  bl_td <- inner_join(bl_td, td1, by = 'stage_mode')

  # format table
  td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)

  # print to html
  cat(paste("Yearly distance (km) travelled by the entire population with ages considered in the model for ", city))
  print(td1)

  cat("\n")

### Baseline total distance for all cities (per year in km)

Baseline distance travelled by entire population with ages considered in the model for all cities

```{r trip_dist_all_cities = "asis", message=F}

# create baseline yearly distance table for the entire population with ages considered in the model for all cities

backup <- bl_td

# add row totals
td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')

# remove all spaces/blanks
for (i in 2:ncol(td)){
  td[, i] <- comprss(td[, i])
}

print(kable(td, caption = paste("Yearly baseline distance table (km) for population with ages considered in the model")))

## Distance by distance category
### Tables

```{r trip_dist_mode_figs = "asis"}
# for each city create a table showing the total distance, the number of trips and the proportion of trips in each distance category

for (city in cities){ # loop through cities

  # extract trip data for the baseline scenario and calculate the total distance, the number of trips and the proportion of trips in each distance category
  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline' | scenario == 'baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% 
    summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), proportion = round(n_vals / nrow(.) * 100, 1))

 # print to html
  print(kable(df, caption = paste("Distance in travel survey by mode for the ages considered in the model for  ", city)))
}

## Daily population distances (per day in km) for ages considered in model
### Tables

```{r trip_dist_mode_figs = "asis"}

for (city in cities){ # loop through distances

  # extract total daily distances travelled by the entire population with ages considered in the model
  df <- io[[city]]$true_dist

  names(df)[-1] <- scen_names_updated

  print(kable(df, caption = paste("Total population distance with ages considered in model by mode per day for  ", city)))
}

@@

### Baseline average duration (mins) per day for all cities

```{r}
# Average baseline duration travelled on each mode for all cities for all people in synthetic population

# calculate the row sum
trip_dur <- trip_dur %>% janitor::adorn_totals('row')

kable(trip_dur, caption = "Avg. duration (min) per person in the synthetic population per day across all cities")

## Health outcomes

### Change in YLLs compared to reference scenario {#change_YLL}

cat(paste0('Reference scenario:  ', reference_scenario))

# Calculate the changes in YLL compared to the reference scenario by age groups

# find ap ylls columns
ap_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))

# find pa yll columns
pa_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))

# # find scenario names of health burden as the reference scenario is not necessarily the baseline scenario
# hb_columns <- names(io[[cities[1]]]$outcomes$hb$ylls)[3: length(names(io[[cities[1]]]$outcomes$hb$ylls))]
# 
# # find the unique scenario names in the hb columns
# inj_scen_cols <- grep("inj", hb_columns) # find location of injury columns
# # extract unique scenario names
# hb_scen_names <- unique(sapply(stringr::str_split(hb_columns[inj_scen_names], "_yll_inj", n = 3), function(x) x[1]))
# # update scenario names
# hb_scen_names_upd <- unlist(lapply(hb_scen_names, FUN=get_qualified_scen_name))

# from the known scenarios, remove the reference scenario
all_orig_scen_upd <- c('Baseline',orig_scen_names_updated)
hb_scen_names_upd <- all_orig_scen_upd[all_orig_scen_upd!=reference_scenario]

# calculate ylls by age category and scenario
# i.e. sum over all disease and remove sex 
yll_totals <- lapply(io[-c(2,3)],function(x){
  temp <- sapply(1:NSCEN,function(y){# loop through scenarios
    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
    xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)]) # sum over all diseases
    sapply(sort(unique(xx$age_cat)),function(z)
      sum(xxx[xx$age_cat == z])) # sum age categories ignoring sex
  })
  colnames(temp) <- hb_scen_names_upd # update scenario names
  temp
})

# add the column total, round the results and print to html
sum_and_round_and_print(yll_totals,"Change total in YLLs compared to the reference scenario in ")
```

### Change in YLLs compared to reference scenario per 100,000 people {#change_YLL_age_100k}

YLLs per 100,000 people by age group by scenario

```{r scen_prop = "asis"}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate change in YLLs compared to the reference scenario per 100,000 people

# extract population statistics
pop_by_age <- lapply(io[-c(2,3)], function(x)
  sapply(unique(x$demographic$age), 
         function(y)
           sum(subset(x$demographic, age == y)$population)
         )
  )

# calculate ylls per 100,000 people
yll_rates <- lapply(cities, function(x)
  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000, # divide by population
        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), # calculate rate for total YLLs
                                             length = NSCEN)*100000
        )
  )

names(yll_rates) <- cities

# round results and print to html
round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
```

### Change in YLLs compared to reference scenario per 100,000 people by age and sex  {#change_YLL_gender_age_100k}

Change in YLLs per 100,000 people by age and sex category for each scenario compared to the reference scenario

cat(paste0('Reference scenario:  ', reference_scenario))

# calculate changes in YLLs for each scenario compared to the Baseline per 100,000 people by age and sex category

# calculate the total ylls for each scenario summing across all diseases
yll_totals <- lapply(io[-c(2,3)], function(x) {
  temp <- sapply(1:NSCEN,function(y){
    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
    rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
  })
  rownames(temp) <- apply(x$outcomes$hb$deaths[,c('sex','age_cat')], 1,
                          function(z)paste0(z[2], '_', z[1]))
  colnames(temp) <- orig_scen_names_updated
  temp
})

# calculate the YLLs per 100,000 people
yll_rates <- lapply(cities, function(x)
  rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')], 1,
                                    function(z) paste0(z[2],'_',z[1])), 
                              rownames(yll_totals[[x]])),] /
          t(repmat(io[[x]]$demographic$population,NSCEN,1)) * 100000,
        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),
                                             length = NSCEN) * 100000))

names(yll_rates) <- cities

# round and print ylls per 100,000 people to html
round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
```

### Change in YLLs due to injury compared to the reference scenario  {#change_yll_injury}

```{r}

cat(paste0('Reference scenario:  ', reference_scenario))

# calculate change in YLLs due to injury

# calculate change in YLLs due to injury by summing across all age categories for the injury results
injury_totals <- lapply(io[-c(2,3)], function(x) {
  xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
  injury_col <- grep("inj", names(xx))[1] 
  xxx <- sapply(1:NSCEN, function(y)
    sapply(sort(unique(xx$age_cat)), function(z)
      sum(xx[xx$age_cat == z, injury_col - 1 + y])))
  colnames(xxx) <- hb_scen_names_upd
  xxx
})

# calculate the total, round the results and print to html
sum_and_round_and_print(injury_totals, "Change in YLLs due to injury in ")
```

### Change in YLLs due to PA compared to the reference scenario {#change_yll_PA}

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLs compared to the reference scenario caused by changes in PA for the unique age categories

# sum the YLLs for each scenario by age category
pa_totals <- lapply(io[-c(2,3)],function(x){
  xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
  xxx <- sapply(1:NSCEN, function(y){
    xxx <- rowSums(xx[, pa_yll_cols[seq(y, length(pa_yll_cols), by = NSCEN)]])
    sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
  })
  colnames(xxx) <- hb_scen_names_upd
  xxx
})

# calculate column sums, round the results and print to html
sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
```

### Change in YLLs due to AP compared to the reference scenario {#change_yll_AP}
```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLS compared to the reference scenario due to changes in AP

# sum the YLLs by scenario and age category
ap_totals <- lapply(io[-c(2,3)],function(x){
  xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
  xxx <- sapply(1:NSCEN,function(y) {
    xxx <- rowSums(xx[, ap_yll_cols[seq(y, length(ap_yll_cols), by = NSCEN)]])
    sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
  })
  colnames(xxx) <- hb_scen_names_upd
  xxx
})

# calculate column sums, round the results and print to html
sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ")
```

### Change in YLLs due to injury compared to the reference scenario per 100,000 people {#change_yll_injury_100k}

<!-- YLLs per 100,000 people by age group by scenario - injury -->

```{r scen_prop = "asis"}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to injury
injury_rates <- lapply(cities, function(x)
  rbind(
    injury_totals[[x]]/t(repmat(pop_by_age[[x]],NSCEN,1)) * 100000,
    Total = colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),
                                            length = NSCEN) * 100000)
  )
names(injury_rates) <- cities

# round results and print to html
round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 people in ")
```

### Change in YLLs due to PA compared to the reference scenario per 100,000 people {#change_yll_PA_100k}

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to PA
pa_rates <- lapply(cities, function(x)
  rbind(
    pa_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN, 1)) * 100000,
    Total = colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),
                                        length = NSCEN) * 100000))
names(pa_rates) <- cities

# round results and print to html
round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 people in ")
```

### Change in YLLs due to AP compared to the reference scenario per 100,000 people {#change_yll_AP_100k}

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to AP
ap_rates <- lapply(cities, function(x)
  rbind(
    ap_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN ,1)) * 100000,
    Total = colSums(ap_totals[[x]]) / rep(sum(pop_by_age[[x]]),
                                          length = NSCEN) * 100000))
names(ap_rates) <- cities

# round results and print to html
round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 people in ")
```

## BY DISEASE

Change in total deaths compared to the reference scenario (for the city based on real population size but only for ages considered in the model) by scenario

### Change in deaths by disease compared to the reference scenario {#change_death_disease}

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the changes in deaths compared to the reference scenario by disease

# sum the total deaths by scenario
disease_totals <- lapply(io[-c(2,3)],function(x){
  # extract disease counts
  xx <- x$outcomes$hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) 
  xxx <- sapply(1:NSCEN,function(y){ # sum over scenarios
    colSums(xx[, seq(y + 2, ncol(xx), by = NSCEN)])
  })
  colnames(xxx) <- hb_scen_names_upd

  # find scenario name of row names
  row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2]))

  # update row names
  rownames(xxx) <- gsub(paste0("sc_",row_scen,"_deaths_"), "", rownames(xxx))

  xxx
})

# sum, round and print the results 
sum_and_round_and_print(disease_totals,"Change in deaths by disease in ")
```

### Change in deaths by disease compared to the reference scenario per 100,000 people  {#change_death_disease_100k}

Change in deaths per 100,000 people (based on real population size but only for ages considered in the model) by scenario in comparison to the reference scenario

```{r scen_prop = "asis"}
cat(paste0('Reference scenario:  ', reference_scenario))

# calculate the change in deaths compared to the reference scenario per 100,000 people

# divided total number of deaths by population
disease_rates <- lapply(cities, function(x)
  rbind(
    disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000,
    Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]), 
                                               length = NSCEN)*100000))
names(disease_rates) <- cities

# round results and print to html
round_and_print(disease_rates,"Change in deaths by disease per 100,000 people in ")
```

### Change in total YLLs by disease compared to the reference scenario {#change_yll_disease}

Change in total YLLs by disease compared to the reference scenario (based on real population size, for ages considered within the ITHIM model)

```{r}
cat(paste0('Reference scenario:  ', reference_scenario))

# sum the total ylls by scenario
disease_totals <- lapply(io[-c(2,3)],function(x){
  xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
  xxx <- sapply(1:NSCEN,function(y) {
    colSums(xx[,seq(y + 2, ncol(xx), by = NSCEN)])
  })
  colnames(xxx) <- hb_scen_names_upd

  # find scenario name of row names
  row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2]))

  # update row names
  rownames(xxx) <- gsub(paste0("sc_",row_scen,"_ylls_"), "", rownames(xxx))
  xxx
})

# sum and round results, print to html
sum_and_round_and_print(disease_totals,"Change in YLL by disease in ")
```

>>>>

## Distance categories

@@

<!-- ### Baseline mode split for all cities (%) -->
<!-- ```{r} -->
<!-- # create baseline mode share table containing information for all cities -->

<!-- # prep data -->
<!-- trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>% -->
<!--   mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>% -->
<!--   janitor::adorn_totals('row') -->

<!-- # print table to html -->
<!-- kable(trip_prop, caption = 'Baseline mode share (%) for all cities - only selected modes') -->

<!-- ``` -->

<!-- ```{r unload_tidyverse, echo = F, message = F} -->
<!-- suppressWarnings({ -->
<!--   detach("package:tidyverse", character.only = T) -->
<!-- }) -->

<!-- ``` -->

@@

<!-- ### Distance table showing average baseline distances in km travelled by each person in the synthetic population by mode for all cities -->

<!-- ```{r} -->
<!-- # create table showing average baseline distances by mode for all cities -->
<!-- trip_dist <- trip_dist %>% janitor::adorn_totals('row') -->

<!-- kable(trip_dist, caption = "Avg. baseline distance (km) per person in the synthetic population per day across all cities") -->

<!-- ``` -->

<!-- ## Distance tables scaled by using total model age population (per year in km) -->

<!-- City specific distance tables for baseline and scenarios for the entire population with ages considered in the model -->

<!-- ```{r trip_tot_dist = "asis", message=F} -->
<!-- # create distance tables for the yearly distances travelled by the entire population -->
<!-- bl_td <- list() -->

<!-- for (city in cities) { # loop through cities -->

<!--   # extract daily population distances for ages considered in model and multiply by 365 to to get yearly distances -->
<!--   td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% -->
<!--     mutate_if(is.numeric, list(~round((.) * 365, 3))) -->

<!--   # update scenario names -->
<!--   colnames(td)[2:ncol(td)] <- scen_names_updated -->

<!--   # match with pre-defined mode order -->
<!--   x <- match(td$stage_mode, trip_order$mode) -->
<!--   x[!is.na(x)] -->
<!--   td <- td[order(x),] -->
<!--   row.names(td) <- NULL -->

<!--   # initialise bl_td with stage_mode -->
<!--   if (length(bl_td) == 0){ -->
<!--     bl_td <- td %>% dplyr::select(stage_mode) -->
<!--   } -->

<!--   # calculate row and column totals -->
<!--   td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col')) -->

<!--   # extract baseline distances -->
<!--   td1 <- td %>% dplyr::select(stage_mode, Baseline) -->

<!--   names(td1)[2] <- city -->

<!--   # join baseline distances for all cities -->
<!--   bl_td <- inner_join(bl_td, td1, by = 'stage_mode') -->

<!--   # format table -->
<!--   td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T) -->

<!--   # print to html -->
<!--   cat(paste("Yearly distance (km) travelled by the entire population with ages considered in the model for ", city)) -->
<!--   print(td1) -->

<!--   cat("\n") -->

<!-- } -->

<!-- ``` -->

<!-- ### Baseline total distance for all cities (per year in km) -->

<!-- Baseline distance travelled by entire population with ages considered in the model for all cities -->

<!-- ```{r trip_dist_all_cities = "asis", message=F} -->

<!-- # create baseline yearly distance table for the entire population with ages considered in the model for all cities -->

<!-- backup <- bl_td -->

<!-- # add row totals -->
<!-- td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row') -->

<!-- # remove all spaces/blanks -->
<!-- for (i in 2:ncol(td)){ -->
<!--   td[, i] <- comprss(td[, i]) -->
<!-- } -->

<!-- print(kable(td, caption = paste("Yearly baseline distance table (km) for population with ages considered in the model"))) -->

<!-- ``` -->

<!-- ## Distance by distance category -->
<!-- ### Tables -->

<!-- ```{r trip_dist_mode_figs = "asis"} -->
<!-- # for each city create a table showing the total distance, the number of trips and the proportion of trips in each distance category -->

<!-- for (city in cities){ # loop through cities -->

<!--   # extract trip data for the baseline scenario and calculate the total distance, the number of trips and the proportion of trips in each distance category -->
<!--   df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline' | scenario == 'baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>%  -->
<!--     summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), proportion = round(n_vals / nrow(.) * 100, 1)) -->

<!--  # print to html -->
<!--   print(kable(df, caption = paste("Distance in travel survey by mode for the ages considered in the model for  ", city))) -->
<!-- } -->

<!-- ``` -->

<!-- ## Daily population distances (per day in km) for ages considered in model -->
<!-- ### Tables -->

<!-- ```{r trip_dist_mode_figs = "asis"} -->

<!-- for (city in cities){ # loop through distances -->

<!--   # extract total daily distances travelled by the entire population with ages considered in the model -->
<!--   df <- io[[city]]$true_dist -->

<!--   names(df)[-1] <- scen_names_updated -->

<!--   print(kable(df, caption = paste("Total population distance with ages considered in model by mode per day for  ", city))) -->
<!-- } -->

<!-- ``` -->

@@

<!-- ### Baseline average duration (mins) per day for all cities -->

<!-- ```{r} -->
<!-- # Average baseline duration travelled on each mode for all cities for all people in synthetic population -->

<!-- # calculate the row sum -->
<!-- trip_dur <- trip_dur %>% janitor::adorn_totals('row') -->

<!-- kable(trip_dur, caption = "Avg. duration (min) per person in the synthetic population per day across all cities") -->

<!-- ``` -->

<!-- ## Health outcomes -->

<!-- ### Change in YLLs compared to reference scenario {#change_YLL} -->

<!-- ```{r scen_prop = "asis"} -->

<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # Calculate the changes in YLL compared to the reference scenario by age groups -->

<!-- # find ap ylls columns -->
<!-- ap_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x)))) -->

<!-- # find pa yll columns -->
<!-- pa_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x)))) -->

<!-- # # find scenario names of health burden as the reference scenario is not necessarily the baseline scenario -->
<!-- # hb_columns <- names(io[[cities[1]]]$outcomes$hb$ylls)[3: length(names(io[[cities[1]]]$outcomes$hb$ylls))] -->
<!-- #  -->
<!-- # # find the unique scenario names in the hb columns -->
<!-- # inj_scen_cols <- grep("inj", hb_columns) # find location of injury columns -->
<!-- # # extract unique scenario names -->
<!-- # hb_scen_names <- unique(sapply(stringr::str_split(hb_columns[inj_scen_names], "_yll_inj", n = 3), function(x) x[1])) -->
<!-- # # update scenario names -->
<!-- # hb_scen_names_upd <- unlist(lapply(hb_scen_names, FUN=get_qualified_scen_name)) -->

<!-- # from the known scenarios, remove the reference scenario -->
<!-- all_orig_scen_upd <- c('Baseline',orig_scen_names_updated) -->
<!-- hb_scen_names_upd <- all_orig_scen_upd[all_orig_scen_upd!=reference_scenario] -->

<!-- # calculate ylls by age category and scenario -->
<!-- # i.e. sum over all disease and remove sex  -->
<!-- yll_totals <- lapply(io[-c(2,3)],function(x){ -->
<!--   temp <- sapply(1:NSCEN,function(y){# loop through scenarios -->
<!--     xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
<!--     xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)]) # sum over all diseases -->
<!--     sapply(sort(unique(xx$age_cat)),function(z) -->
<!--       sum(xxx[xx$age_cat == z])) # sum age categories ignoring sex -->
<!--   }) -->
<!--   colnames(temp) <- hb_scen_names_upd # update scenario names -->
<!--   temp -->
<!-- }) -->

<!-- # add the column total, round the results and print to html -->
<!-- sum_and_round_and_print(yll_totals,"Change total in YLLs compared to the reference scenario in ") -->
<!-- ``` -->

<!-- ### Change in YLLs compared to reference scenario per 100,000 people {#change_YLL_age_100k} -->

<!-- YLLs per 100,000 people by age group by scenario -->

<!-- ```{r scen_prop = "asis"} -->
<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # calculate change in YLLs compared to the reference scenario per 100,000 people -->

<!-- # extract population statistics -->
<!-- pop_by_age <- lapply(io[-c(2,3)], function(x) -->
<!--   sapply(unique(x$demographic$age),  -->
<!--          function(y) -->
<!--            sum(subset(x$demographic, age == y)$population) -->
<!--          ) -->
<!--   ) -->

<!-- # calculate ylls per 100,000 people -->
<!-- yll_rates <- lapply(cities, function(x) -->
<!--   rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000, # divide by population -->
<!--         Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), # calculate rate for total YLLs -->
<!--                                              length = NSCEN)*100000 -->
<!--         ) -->
<!--   ) -->

<!-- names(yll_rates) <- cities -->

<!-- # round results and print to html -->
<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
<!-- ``` -->

<!-- ### Change in YLLs compared to reference scenario per 100,000 people by age and sex  {#change_YLL_gender_age_100k} -->

<!-- Change in YLLs per 100,000 people by age and sex category for each scenario compared to the reference scenario -->

<!-- ```{r scen_prop = "asis"} -->

<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # calculate changes in YLLs for each scenario compared to the Baseline per 100,000 people by age and sex category -->

<!-- # calculate the total ylls for each scenario summing across all diseases -->
<!-- yll_totals <- lapply(io[-c(2,3)], function(x) { -->
<!--   temp <- sapply(1:NSCEN,function(y){ -->
<!--     xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
<!--     rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)]) -->
<!--   }) -->
<!--   rownames(temp) <- apply(x$outcomes$hb$deaths[,c('sex','age_cat')], 1, -->
<!--                           function(z)paste0(z[2], '_', z[1])) -->
<!--   colnames(temp) <- orig_scen_names_updated -->
<!--   temp -->
<!-- }) -->

<!-- # calculate the YLLs per 100,000 people -->
<!-- yll_rates <- lapply(cities, function(x) -->
<!--   rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')], 1, -->
<!--                                     function(z) paste0(z[2],'_',z[1])),  -->
<!--                               rownames(yll_totals[[x]])),] / -->
<!--           t(repmat(io[[x]]$demographic$population,NSCEN,1)) * 100000, -->
<!--         Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), -->
<!--                                              length = NSCEN) * 100000)) -->

<!-- names(yll_rates) <- cities -->

<!-- # round and print ylls per 100,000 people to html -->
<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to injury compared to the reference scenario  {#change_yll_injury} -->

<!-- ```{r} -->

<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # calculate change in YLLs due to injury -->

<!-- # calculate change in YLLs due to injury by summing across all age categories for the injury results -->
<!-- injury_totals <- lapply(io[-c(2,3)], function(x) { -->
<!--   xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
<!--   injury_col <- grep("inj", names(xx))[1]  -->
<!--   xxx <- sapply(1:NSCEN, function(y) -->
<!--     sapply(sort(unique(xx$age_cat)), function(z) -->
<!--       sum(xx[xx$age_cat == z, injury_col - 1 + y]))) -->
<!--   colnames(xxx) <- hb_scen_names_upd -->
<!--   xxx -->
<!-- }) -->

<!-- # calculate the total, round the results and print to html -->
<!-- sum_and_round_and_print(injury_totals, "Change in YLLs due to injury in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to PA compared to the reference scenario {#change_yll_PA} -->

<!-- ```{r} -->
<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # calculate the changes in YLLs compared to the reference scenario caused by changes in PA for the unique age categories -->

<!-- # sum the YLLs for each scenario by age category -->
<!-- pa_totals <- lapply(io[-c(2,3)],function(x){ -->
<!--   xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
<!--   xxx <- sapply(1:NSCEN, function(y){ -->
<!--     xxx <- rowSums(xx[, pa_yll_cols[seq(y, length(pa_yll_cols), by = NSCEN)]]) -->
<!--     sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z])) -->
<!--   }) -->
<!--   colnames(xxx) <- hb_scen_names_upd -->
<!--   xxx -->
<!-- }) -->

<!-- # calculate column sums, round the results and print to html -->
<!-- sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to AP compared to the reference scenario {#change_yll_AP} -->
<!-- ```{r} -->
<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # calculate the changes in YLLS compared to the reference scenario due to changes in AP -->

<!-- # sum the YLLs by scenario and age category -->
<!-- ap_totals <- lapply(io[-c(2,3)],function(x){ -->
<!--   xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
<!--   xxx <- sapply(1:NSCEN,function(y) { -->
<!--     xxx <- rowSums(xx[, ap_yll_cols[seq(y, length(ap_yll_cols), by = NSCEN)]]) -->
<!--     sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z])) -->
<!--   }) -->
<!--   colnames(xxx) <- hb_scen_names_upd -->
<!--   xxx -->
<!-- }) -->

<!-- # calculate column sums, round the results and print to html -->
<!-- sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to injury compared to the reference scenario per 100,000 people {#change_yll_injury_100k} -->

<!-- <!-- YLLs per 100,000 people by age group by scenario - injury --> -->

<!-- ```{r scen_prop = "asis"} -->
<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # calculate the changes in YLLs compared to the reference scenario per 100,000 people due to injury -->
<!-- injury_rates <- lapply(cities, function(x) -->
<!--   rbind( -->
<!--     injury_totals[[x]]/t(repmat(pop_by_age[[x]],NSCEN,1)) * 100000, -->
<!--     Total = colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]), -->
<!--                                             length = NSCEN) * 100000) -->
<!--   ) -->
<!-- names(injury_rates) <- cities -->

<!-- # round results and print to html -->
<!-- round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 people in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to PA compared to the reference scenario per 100,000 people {#change_yll_PA_100k} -->

<!-- ```{r} -->
<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # calculate the changes in YLLs compared to the reference scenario per 100,000 people due to PA -->
<!-- pa_rates <- lapply(cities, function(x) -->
<!--   rbind( -->
<!--     pa_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN, 1)) * 100000, -->
<!--     Total = colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]), -->
<!--                                         length = NSCEN) * 100000)) -->
<!-- names(pa_rates) <- cities -->

<!-- # round results and print to html -->
<!-- round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 people in ") -->
<!-- ``` -->

<!-- ### Change in YLLs due to AP compared to the reference scenario per 100,000 people {#change_yll_AP_100k} -->

<!-- ```{r} -->
<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->

<!-- # calculate the changes in YLLs compared to the reference scenario per 100,000 people due to AP -->
<!-- ap_rates <- lapply(cities, function(x) -->
<!--   rbind( -->
<!--     ap_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN ,1)) * 100000, -->
<!--     Total = colSums(ap_totals[[x]]) / rep(sum(pop_by_age[[x]]), -->
<!--                                           length = NSCEN) * 100000)) -->
<!-- names(ap_rates) <- cities -->

<!-- # round results and print to html -->
<!-- round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 people in ") -->
<!-- ``` -->

<!-- ## BY DISEASE -->

<!-- Change in total deaths compared to the reference scenario (for the city based on real population size but only for ages considered in the model) by scenario -->

<!-- ### Change in deaths by disease compared to the reference scenario {#change_death_disease} -->
